\bf On semidefinite-representable sets over valuation fields

### 2024-11-11 article/paper_padic_spec.tex
commit by Simone

\begin{definition}
  We say that a matrix $M \in \allmat_{d,d}(K)$ is \emph{positive semidefinite} (in symbols $M \succeq 0$)
  if all its eigenvalues have nonnegative valuation in an algebraic closure of $K$.
  We denote by $\allmat_{d,d}^+(K) \subset \allmat_{d,d}(K)$ the set of positive semidefinite matrices.
  The characteristic polynomial of a matrix $M \in \allmat_{d,d}(K)$ is denoted $\chi_M \in K[T]$.
\end{definition}

@@

\begin{theorem}
  \label{caracsdp}
  $M \in \allmat_{d,d}^+(K)$ if and only if $\chi_M \in \OK[T]$.
\end{theorem}
\begin{proof}
  Both senses of the equivalence will be proven by contraposition.

  Let $M \in \allmat_{d,d}(K)$ with characteristic polynomial $\chi_M = T^d + \sum_{i=0}^{d-1} a_i T^i$ for $a_0,...,a_{d-1} \in K$.
  Firstly, suppose $M \not \in \allmat_{d,d}^+(K)$, {\it i.e.} $\chi_M$ has a root $\rho$ with $\val\,\rho<0$.
  It follows that
  $$
  \val \chi_M (\rho) = \val \left(\rho^d+ \sum_{i=0}^{d-1} a_i \rho^i\right) = + \infty
  $$ and $\val \rho^d = d \val \rho <0$.
  The properties of the non-archimedean inequality and of the valuation imply that
  $$
  \min_{0 \leq i \leq d-1} \val \left(a_i \rho^i\right) \leq \val\left(\sum_{i=0}^{d-1} a_i \rho^i\right) = \val \rho^d
  $$
  %%%$\val ( \sum_{i=0}^{d-1} a_i \rho^i) = \val \rho^d$ and $\val ( \sum_{i=0}^{d-1} a_i \rho^i) \ge \min_{i=0,..., d-1} \val \left(a_i \rho^i\right)$.
  Therefore, there exists $j \in \left\{0,...,d-1\right\}$ such that $\val(a_j \rho^j) \le \val\rho^d$,
  {\it i.e.} such that $\val a_j \le (d-j) \val \rho <0$. This shows that $\chi_M \not\in \OK[T]$.

>>>>

\begin{definition}
  We say that a matrix $M \in \allmat_{d,d}(K)$ is \emph{positive semidefinite} ($M \succeq 0$)
  if all its eigenvalues have nonnegative valuation in an algebraic closure of $K$.
  We denote by $\allmat_{d,d}^+(K) \subset \allmat_{d,d}(K)$ the set of positive semidefinite matrices.
  The characteristic polynomial of a matrix $M \in \allmat_{d,d}(K)$ is denoted $\chi_M \in K[T]$.
\end{definition}

@@

Testing positive semidefiniteness reduces to checking the coefficients of $\chi_M$.
\begin{theorem}
  \label{caracsdp}
  $M \in \allmat_{d,d}^+(K)$ if and only if $\chi_M \in \OK[T]$.
\end{theorem}
\begin{proof}
  Both senses of the equivalence will be proven by contraposition.
  Let $M \in \allmat_{d,d}(K)$ with characteristic polynomial $\chi_M = T^d + \sum_{i=0}^{d-1} a_i T^i$ for $a_0,...,a_{d-1} \in K$.
  Firstly, suppose $M \not \in \allmat_{d,d}^+(K)$, {\it i.e.} $\chi_M$ has a root $\rho$ with $\val\,\rho<0$.
  It follows that
  $$
  \val \chi_M (\rho) = \val \left(\rho^d+ \sum_{i=0}^{d-1} a_i \rho^i\right) = + \infty
  $$ and $\val \rho^d = d \val \rho <0$.
  The properties of the non-archimedean inequality and of the valuation imply that
  $$
  \min_{0 \leq i \leq d-1} \val \left(a_i \rho^i\right) \leq \val\left(\sum_{i=0}^{d-1} a_i \rho^i\right) = \val \rho^d
  $$
  %%%$\val ( \sum_{i=0}^{d-1} a_i \rho^i) = \val \rho^d$ and $\val ( \sum_{i=0}^{d-1} a_i \rho^i) \ge \min_{i=0,..., d-1} \val \left(a_i \rho^i\right)$.
  Therefore, there exists $j \in \left\{0,...,d-1\right\}$ such that $\val(a_j \rho^j) \le \val\rho^d$,
  {\it i.e.} such that $\val a_j \le (d-j) \val \rho <0$. This shows that $\chi_M \not\in \OK[T]$.
